---
title: "data_analysis"
author: "Chenyu Li"
format: html
editor: source
editor_options: 
  chunk_output_type: console
---

# 1 Initial setting

## 1.1 clear workspacce and set default color

```{r reset, include=FALSE}
graphics.off()
rm(list=ls(all.names=TRUE))
options(digits = 3)
```

## 1.2 import functions and packages

```{r}

# package list
packages = c("tidyverse", "afex", "emmeans")
# get the package names that are not installed
missing_packages <- setdiff(packages, rownames(installed.packages()))
# install the missing packages
if (length(missing_packages) > 0) install.packages(missing_packages)
suppressMessages(sapply(packages, library, character.only = TRUE))

# import private functions
function_path <- "./functions"
function_files <- list.files(function_path, pattern = "*.R", full.names = TRUE)
sapply(function_files, source)

# theme
theme_set(theme_bw())
dodge1 = position_dodge(.1)
dodge2 = position_dodge(.2)
dodge4 = position_dodge(.4)


# task information
groupID <- "group5"
IV_list = list(
  group0 = c("setsize", "layout"),
  group1 = c("cueType"),
  group2 = c("radius", "setsize"),
  group3 = c("nTarget", "nDistractor"),
  group4 = c("ratio", "layout"),
  group5 = c("colorCon", "layout")
)
IVs <- IV_list[[groupID]]
sample_path <- str_glue("./models/sample_{groupID}/")
figure_path <- str_glue("./figures/figures_{groupID}/")
# check whether the folders are existent or not. If not, create a new one
dir.create(file.path(sample_path), showWarnings = FALSE)
dir.create(file.path(figure_path), showWarnings = FALSE)

```


# 2 Data processing

## 2.1 Load data

```{r}

# behavior data
Raw_Res <- str_glue("./data/Expra_{groupID}_Res.csv") %>% # the name of the file
  read_csv() %>% 
  filter(screenID == "judge") 

# survey data
Data_survey <- str_glue("./data/Expra_{groupID}_survey.csv") %>% 
  read_csv()

```

## 2.2 Exclude participants
1. Exclude participants who were distracted during the task

```{r}

distractor_list <- Data_survey %>%
  filter(str_detect(attention, "Ja|Yes")) %>% 
  pull(participant) 

low_acc_list <- Raw_Res %>% 
  summarise(
    acc = mean(acc, na.rm = TRUE),
    .by = c(participant)
  ) %>% 
  filter(acc < 0.7) %>% 
  pull(participant)

excluding_list <- c(distractor_list, low_acc_list)

Data_Res <- Raw_Res %>% 
  filter(!participant %in% excluding_list)

consistent_list <- Data_Res %>% 
  filter(colorCon == "fixed") %>% 
  select(participant)

switch_list <- Data_Res %>% 
  filter(colorCon == "switch") %>% 
  select(participant)

random_list <- Data_Res %>% 
  filter(colorCon == "random") %>% 
  select(participant)


str_glue(
  "{length(excluding_list)} participants were excluded from the analysis, ",
  "leaving {n_distinct(Data_Res$participant)} for the data analysis") %>% message()

str_glue(
  "Consistent: {n_distinct(consistent_list$participant)} participants\n",
  "Switch    : {n_distinct(switch_list$participant)} participants\n",
  "Random    : {n_distinct(random_list$participant)} participants"
) %>% message()

```

## 2.3 Exclude trials
1. Exclude trials with RT < 100 ms or > 10000 ms
2. Exclude trials with incorrect responses
3. Exclude trials with RT > 3 SD from the mean RT of each participant in each condition

```{r}

Data_ACC <- Data_Res
  
Data_RT <- Data_Res %>% 
  filter(RT > 100, RT < 10000, acc == 1) %>%
  mutate(zRT = (RT - mean(RT))/sd(RT),
         .by = c(participant,condition)) %>%
  filter(abs(zRT) < 3) %>%
  mutate(RT = RT/1000) # convert RT to seconds

```

# 3 Data visualization

## 3.1 Aggregate data
1. Dependent variable: accuracy, reaction time
2. Independent variables:
  - within-subject: condition, set size, layout
  - between-subject: ?

```{r}

Table_ACC <- Data_ACC %>% 
  summarise(
    mean = mean(acc),
    sd = sd(acc),
    se = sd/sqrt(n()),
    ci = 1.96 * se,
    .by = c(condition, all_of(IVs))
  ) %>% 
  labelConditions()

Table_RT <- Data_RT %>% 
  summarise(
    mean = mean(RT),
    sd = sd(RT),
    se = sd/sqrt(n()),
    ci = 1.96 * se,
    .by = c(condition, all_of(IVs))
  ) %>% 
  labelConditions()


Data_RT_Benefits <- Data_RT %>% 
  summarise(
    RT = mean(RT), # calculate the mean RT
    .by = c(participant, condition, all_of(IVs))
  ) %>% 
  pivot_wider(names_from = condition, values_from = RT) %>% 
  mutate(
    PB = neutral - positive, 
    NB = neutral - negative 
  )

Table_RT_Benefits <- Data_RT_Benefits %>%
  select(participant, all_of(IVs), PB, NB) %>% 
  pivot_longer(cols = c(PB, NB), names_to = "benefit", values_to = "RT") %>% 
  summarise(
    mean = mean(RT, na.rm = TRUE),
    sd = sd(RT, na.rm = TRUE),
    se = sd/sqrt(n()),
    ci = 1.96 * se,
    .by = c(benefit, all_of(IVs))
  ) %>% 
  labelConditions()

```

## 3.2 Plot results

### 3.2.1 Accuracy

```{r}

Plot_ACC <- Table_ACC %>% 
  ggplot(aes(x = colorCon, y = mean, group = condition, color = condition)) +
  geom_line(position = dodge2)+
  geom_point(position = dodge2)+
  geom_errorbar(aes(ymin = mean - ci, ymax = mean + ci), width = .1, position = dodge2) +
  facet_wrap(~layout)+
  ylim(0.5,1.05)+
  labs(
    x = "Group",
    y = "Proportion of correct",
    color = "Condition"
  )+
  theme(
    # legend.position = "none",
    axis.text = element_text(size = 10.5),
    axis.title = element_text(size = 11),
    strip.text = element_text(size = 10.5))

ggsave(
  plot = Plot_ACC,
  path = figure_path,
  filename = "Plot_Acc.pdf",
  width = 8, height = 4
)

```

### 3.2.2 Reaction time (condition)

```{r}

Plot_RT_condition <- Table_RT %>% 
  ggplot(aes(x = colorCon, y = mean, group = condition, color = condition)) +
  geom_line(position = dodge4)+
  geom_point(position = dodge4)+
  geom_errorbar(aes(ymin = mean - ci, ymax = mean + ci), width = .1, position = dodge4) +
  facet_wrap(~layout)+
  ylim(0.8, 3)+
  labs(
    x = "Group",
    y = "Reaction time (s)",
    color = "Condition"
  )+
  theme(
    # legend.position = "none",
    axis.text = element_text(size = 10.5),
    axis.title = element_text(size = 11),
    strip.text = element_text(size = 10.5))

ggsave(
  plot = Plot_RT_condition,
  path = figure_path,
  filename = "Plot_RT_condition.pdf",
  width = 8, height = 4
)

```

### 3.2.3 Reaction time (benefits)

```{r}

Plot_RT_benefits <- Table_RT_Benefits %>% 
  ggplot(aes(x = colorCon, y = mean, group = benefit, color = benefit)) +
  geom_line(position = dodge4)+
  geom_point(position = dodge4)+
  geom_errorbar(aes(ymin = mean - ci, ymax = mean + ci), width = .1, position = dodge4) +
  facet_wrap(~layout)+
  ylim(0, 1)+
  labs(
    x = "Group",
    y = "Reaction time benefit (s)",
    color = "Benefit"
  )+
  theme(
    axis.text = element_text(size = 10.5),
    axis.title = element_text(size = 11),
    strip.text = element_text(size = 10.5))

ggsave(
  plot = Plot_RT_benefits,
  path = figure_path,
  filename = "Plot_RT_Benefits.pdf",
  width = 8, height = 4
)

```


# 4 Data analysis

## 4.1 Aggregated data

```{r}

mData_ACC <- Data_ACC %>% 
  summarise(
    ACC = mean(acc),
    .by = c(participant, condition, all_of(IVs))
  )


mData_RT_condition <- Data_RT %>% 
  summarise(
    RT = mean(RT),
    .by = c(participant, condition, all_of(IVs))
  )


mData_RT_benefits <- mData_RT_condition %>% 
  pivot_wider(names_from = condition, values_from = RT) %>% 
  mutate(PB =  neutral - positive,
         NB =  neutral - negative)


```

## 4.2 Accuracy

```{r}

```


## 4.3 Reaction time

### 4.3.1 difference in RT across conditions

```{r}


```

### 4.3.2 benefit of negative and positive information

```{r}


```

